Data Analysis Methods for the Detection of the Epoch of Reionization

### 2017-11-22 PAPER.tex
made jonnie changes until Section 3

In the next three sections, we investigate the effects of weighting matrices on signal loss by experimenting with different matrices $\textbf{R}$ and examine their impact on the resulting power spectra estimate $\hat{\textbf{p}}$. Our ultimate goal in experimenting with weighting is to suppress foregrounds and investigate EoR losses associated with it.

@@

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.2cm 0.3cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss2.png}
	\caption{Eigenspectrum of $\hat{\textbf{C}}_{FG}$ (blue), $\hat{\textbf{C}}_{EoR}$ (red), and $\hat{\textbf{C}}_{FG+EoR}$ (green). The eigenspectrum of $\hat{\textbf{C}}_{FG}$ peaks at the zeroth eigenmode, due to the presence of only one sinusoid. These empirically estimated covariance matrices have eigenspectra that are different from that of a true $\textbf{C}$, meaning that some eigenmodes (shapes in the data) may be down-weighted more significantly than they should be, producing signal loss.}
	\label{fig:toy_sigloss2}
\end{figure}

If the true covariance matrix $\textbf{C}$ of our data was known, then every single eigenvalue of $\textbf{C}$ would be representative of real fluctuations in the data. However, when using an estimated $\hat{\textbf{C}}$ that is derived from one particular data realization, some of its eigenvalues may be describing random fluctuations that just happen to exist in the data realization. Said differently, shapes that may not exist (or have a weaker existence) in a true covariance may appear stronger in the estimated covariance. Hence, they will be down-weighted more than they should be.

In general, the strongest modes of $\hat{\textbf{C}}$ (highest eigenvalues) are expected to be dominated by bright foregrounds --- the most prominent shapes in the data. If bright foregrounds dominate a dataset, these highest eigenvalue modes would exist in both $\textbf{C}$ and $\hat{\textbf{C}}$. This is demonstrated in the toy model by the successful suppression of the foreground mode, where our estimated covariance matrix identifies the sinusoid shape and assigns it the highest eigenvalue (the peak in Figure \ref{fig:toy_sigloss2}). In Figure \ref{fig:toy_sigloss12} we show the covariances of our toy model datasets along with inverse covariance weighted data. The foreground sinusoid is clearly visible in $\hat{\textbf{C}}_{FG}$ but effectively absent from $\hat{\textbf{C}}^{-1}_{FG}\textbf{x}_{FG}$.

@@

The danger of an empirically estimated covariance matrix comes from not being able to describe weak eigenmodes of $\textbf{C}$ accurately. Instead, the weak eigenmodes of $\hat{\textbf{C}}$ may characterize noise-like fluctuations which will be down-weighted. These noise fluctuations could be the signal itself, since $\hat{\textbf{C}}$ is constructed using data that contains the signal. This leads to the `overfitting' of noise (or `overfitting' of signal), which is signal loss.

Using what we've learned about the eigenspectrum, we can tweak it in a simple way to suppress foregrounds and yield zero signal loss. Recall that our toy model foreground is a sinusoid that can be perfectly described by a single eigenmode. We can project out the zeroth eigenmode (i.e. zero out all eigenmodes except the first one), thereby down-weighting the foregrounds perfectly and nothing else. Hence, we are not using information from all the weaker eigenmodes which carry the risk of describing fluctuations that may not exist in the true covariance. Altering $\hat{\textbf{C}}$ as such is one example of a regularization method, in which we are changing $\hat{\textbf{C}}$ in a way that flattens its eigenspectrum, making it more similar to that of $\textbf{I}$. The resulting power spectrum estimate for this case is shown in the right plot of Figure \ref{fig:toy_sigloss3}. In this case we recover EoR, demonstrating that if we can model our foregrounds perfectly (and disentangle them from other modes), we can down-weight them without signal loss. There are several other ways to regularize $\hat{\textbf{C}}$, and we will discuss some in Section \ref{sec:otherweight}.

@@

We have shown how signal loss can arise due to inaccurately characterizing weak eigenmodes with a data-estimated covariance. We will next show how this effect is exaggerated by reducing the total number of independent samples in a dataset.

@@

\begin{figure*}
	\centering
	\includegraphics[trim={1.5cm 0.2cm 1.5cm 0.3cm},clip,height=0.29\textwidth]{plots/toy_sigloss5.png}
	\includegraphics[trim={0.3cm 0.2cm 0.3cm 0.3cm},clip,height=0.29\textwidth]{plots/toy_sigloss15.png}
	\caption{Left: Our `fringe-rate filtered' (time-averaged) toy model dataset. We average every four samples together, yielding $25$ independent samples in time. Real parts are shown here. Right: Eigenspectrum of $\hat{\textbf{C}}_{FG+EoR}$, in the case of no fringe-rate filtering (solid green) and with fringe-rate filtering (dashed green). The dashed spectrum is steeper, resulting in more dramatic differences in weighting between eigenmodes. Consequently, if all eigenmodes are used in weighting it is possible to down-weight random noisy modes more severely than for the non-averaged case.}
	\label{fig:toy_sigloss5}
\end{figure*}

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.2cm 0.3cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss7.png}
	\caption{Resulting power spectrum estimate for the fringe-rate filtered toy model simulation --- foregrounds only (blue), EoR only (red), and the weighted FG + EoR dataset (green). We use inverse covariance weighting where $\textbf{C}$ is computed from the data. There is a larger amount of signal loss than for the non-averaged data, a consequence of a steepened, weakly characterized eigenspectrum from using fewer independent modes in the data. The power spectrum estimate is also noisier, since most of the information used in constructing the estimate came from only a few eigenmodes.}
	\label{fig:toy_sigloss7}
\end{figure}

@@

Using our toy model, we have seen that a sensitivity-driven analysis technique like fringe-rate filtering has trade-offs of signal loss and noisier estimates when using data-estimated covariance matrices. Longer integrations increase sensitivity but reduce the number of independent samples, resulting in poorly characterized, steep eigenspectra that can overfit noise greatly. We note that a fringe-rate filter does have a range of benefits, many described in \citet{parsons_et_al2016}, so it can still be advantageous to use one despite the trade-offs.

@@

In Section \ref{sec:toymodel} we showed one example (projecting the zeroth eigenmode) of how altering $\hat{\textbf{C}}$ can make the difference between zero and some signal loss, if we can distinguish between real eigenmodes in a true covariance matrix from random fluctuation-induced eigenmodes in an estimated one. We will now use our toy model to describe several other ways to tailor $\hat{\textbf{C}}$ in order to minimize signal loss. We illustrate the resulting power spectra and eigenspectra for four different cases in Figures \ref{fig:toy_sigloss8} and \ref{fig:toy_sigloss14}.

As a first test, we model the covariance matrix of EoR as a proof of concept that if perfect models are known, signal loss can be avoided. We know that our simulated EoR signal should have a covariance matrix that mimics the identity matrix, with its variance encoded along the diagonal. We model $\textbf{C}_{EoR}$ as such, instead of computing it based on $\textbf{x}_{EoR}$ itself. Next, we add $\textbf{C}_{EoR} + \hat{\textbf{C}}_{FG}$ (where $\hat{\textbf{C}}_{FG} = \langle\textbf{x}_{FG}\textbf{x}_{FG}^{\dagger}\rangle$) to obtain a final $\hat{\textbf{C}}$ to use in weighting. In Figure \ref{fig:toy_sigloss8} (upper left), we see that there is negligible signal loss. This is because by modeling $\textbf{C}_{EoR}$, we avoid over-fitting small fluctuations in the data that our model doesn't know about (but an empirically derived $\hat{\textbf{C}}$ would). This is evident when comparing the green and red curves in Figure \ref{fig:toy_sigloss14}. We don't see perfect EoR recovery though, as the green and red curves vary slightly in Figure \ref{fig:toy_sigloss8}. This deviation is a consequence of the difference between the true and modeled covariance matrices. In practice such a weighting option is not feasible, as we do not know $\textbf{C}_{EoR}$.

@@

The second panel (top right) in Figure \ref{fig:toy_sigloss8} uses a regularization method of setting $\hat{\textbf{C}} = \hat{\textbf{C}} + \gamma\textbf{I}$, where $i$$=$$j$$=$$20$ (number of frequencies) and $\gamma = 5$ (an arbitrary strength of $\textbf{I}$ for the purpose of this toy model). By adding the identity matrix, element-wise, we are weighting the diagonal elements of the matrix more heavily than those off-diagonal, thereby flattening out its eigenspectrum. If all modes are given similar weights, we avoid down-weighting artificial modes more than others.

@@

Our second major $21$ cm power spectrum theme is error estimation, as we desire robust methods for determining accurate confidence intervals for our measurements. Two popular ways of empirically calculating errors on a power spectrum measurement are calculating the variance of a dataset, and computing a theoretical error estimate based on an instrument's system temperature and observational parameters. In a perfect world, both methods would match up. However, in practice the two don't always agree due to a number of factors, including time, frequency, and antenna-dependent noise and non-uniform weightings. For PAPER's analysis, we choose to compute error bars on our measurements that have been derived from its inherent variance.

A common technique used to estimate the error in a measurement is bootstrapping. Bootstrapping uses sampling with replacement to estimate a posterior distribution. For example, power-spectral measurements of $21$ cm data can be made from different samples of data (i.e. data can be sliced along time, baselines, etc. and a power spectrum measurement can be made for each). Averaging each sample set gives us a realization of the power spectrum answer. Realizations are correlated with each other to a degree set by the fraction of sampled points that are held in common between them. Hence, through the process of resampling and averaging along these axes, we can estimate error bars for our results which represent the underlying distribution of power spectra values that are allowed by our measurements.

Again, we focus on toy models to highlight pitfalls that one can fall into when bootstrapping power spectra. Suppose we have a Gaussian random signal dataset of length $N=1000$ and unity variance. We are interested in the average of the dataset, and predict that the error on the mean should obey $1/\sqrt{N}$, where $N$ is the number of samples.

We form $100$ bootstraps, each comprised of an array of length $N$ that is created by a random resampling of the original data, with replacement. The standard deviation over the $100$ bootstraps gives an error estimate for our dataset. As shown in Figure \ref{fig:toy_error1} by the grey star, the error computed from bootstrapping matches our theoretical prediction (green).

@@

This example highlights the importance of understanding how analysis techniques (e.g. fringe-rate filtering) can affect a common statistical procedure like bootstrapping. Bootstrapping as a means of estimating power spectrum errors from real fringe-rate filtered data requires knowledge of the number of independent samples, which is not always a trivial task. For example, computing the effective number of independent samples of fringe-rate filtered data is not as simple as counting the number of averages performed. Down-sampling a fringe-rate filtered signal is straightforward using a boxcar average, but non-trivial with a more complicated convolution function that has long tails. Hence, we do not recommend bootstrapping unless the number of independent samples along the axis that is being re-sampled is well-determined.

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.3cm 0.3cm 0.3cm},width=\columnwidth]{plots/toy_error1.png}
	\caption{Error estimation produced by bootstrapping as a function of the number of elements drawn per bootstrap when sampling with replacement. The star represents the error associated with drawing $1000$ elements from a length $1000$ array of a Gaussian random signal. The black points correspond to time-averaged data (correlated data) which has $100$ independent samples. They illustrate how errors can be underestimated if drawing more elements than there are independent samples in the data. The estimated errors match up with the theoretical prediction only at $N=100$.}
	\label{fig:toy_error1}
\end{figure}

We will now discuss a second subtle feature of bootstrapping that can lead to an over-estimation of errors. Suppose we have $5$ independent measurements of the sky (from $5$ different baselines, for example). A bootstrap then consists of $5$ measurements that are drawn randomly with replacement from the original set. If all $5$ spaces are filled randomly, there is a high probability that some measurements will be repeated in the bootstrap because they are drawn more than once. The bootstrap may therefore consist of only $3$ or $4$ independent measurements --- a number smaller than the total number of samples.  In fact, the probability of drawing $5$ completely independent measurements is less than $4\%$.

@@

Therefore, we use a slightly modified bootstrapping method. For each bootstrap, we draw $N-1=4$ samples without replacement and draw the last slot randomly with replacement. In doing so there is a small chance a bootstrap will consist of $5$ independent measurements, but even with $4$ our sensitivity is nearly maximized. One may wonder whether this change is still a legitimate way of error estimating since random sampling only occurs for one value in a bootstrap. However, as long as the number of possible variations is greater than the number of bootstraps we perform, it is a valid way to uncover the inherent variability in a dataset.

@@

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.3cm 0.3cm 0.3cm},width=\columnwidth]{plots/toy_error2.png}
	\caption{The fraction of independent elements in one bootstrap sample as a function of the total number of elements in a dataset (i.e. if $4$ out of $5$ elements are independent, this would yield a fraction of $0.8$). A fraction of $1.0$ means that all elements in the bootstrap sample are independent, and therefore sensitivity is maximized (alternately, this axis can be thought of as power spectrum sensitivity, where sensitivity increases moving upwards along the y-axis). Two bootstrapping methods are shown here --- sampling all elements with replacement (grey) and sampling only the last element with replacement (black). The former fails to maximize sensitivity since some elements end up being repeated due to random sampling. The `outlier' points lying at a fraction of $1.0$ at the far left of the plot occur in the off-chance that the last spot in the sample happens to be the one missing element instead of a repeated element --- this is more likely to happen with a small total number of elements but also results in a more dramatically different fraction as compared to the case of a repeated element in the last spot, hence why these points appear like outliers.}
	\label{fig:toy_error2}
\end{figure}

@@

However, this isolation is not perfect.  In deep measurements, power spectrum measurements at $k_{\parallel}$ values beyond the length of a baseline are often still contaminated at a low level. This leakage can be attributed to convolution kernels associated with Fourier-transforming visibilities into delay-space. In other words, smooth-spectrum foregrounds appear as $\delta$-functions in delay-space, convolved by the Fourier transform of the source spectrum and the antenna response, both of which could smear out the foregrounds and cause leakage outside the wedge.

@@

Confirming foreground detections at higher $k$'s is more difficult. In the next section, we will present some tests that can help distinguish these excesses from that of EoR.

@@

In the next section, we approach the difficult task of tracing excesses to foreground, noise, and EoR biases through a discussion of useful jackknife tests.

@@

The jackknife is a resampling technique in which a statistic (i.e. power spectrum) is computed in subsets of the data. These subsets are then compared to reveal systematics. In this section we define two main tests --- the null test and the traditional jackknife --- and explain how a power spectrum detection must pass each. We then highlight how these tests can be used to help distinguish between different sources of bias.

\begin{itemize}
\item \textbf{Null Test}: A null test is a type of jackknife test that removes the astronomical signal from data in order to investigate underlying systematics. Several null tests on data are described in \citet{keating_et_al2016}. For example, one can divide data into two subsets by separating odd and even julian dates, or the first half of the observing season from the second. Subtracting the two removes signal that is common to both subsets, including foregrounds and EoR. The resulting power spectrum should be consistent with thermal noise estimates; if it is not, it suggests the presence of a systematic that differs from one of the data subsets to the other (i.e. doesn't get subtracted perfectly). 
\item \textbf{Traditional Jackknife}: In a more broad sense, it is important to perform many jackknife tests in order to instill confidence in a final result. A stable result must be steadfast throughout all jackknives no matter how the data is sliced. Jackknives can be taken along several different axes --- for example, one could start with a full dataset, and compute a new power spectrum every time as a day of data is removed, or a baseline is removed. This type of jackknife would reveal bias present only at certain LSTs (such as a foreground source), for example, or misbehaving baselines.
\end{itemize}

@@

\begin{eqnarray}
\label{eq:bias1}
\textbf{x}_{a} &=& \textbf{n}_{a} + \textbf{e} + \textbf{fg} \\
\label{eq:bias2}
\textbf{x}_{b} &=& \textbf{n}_{b} + \textbf{e} + \textbf{fg},
\end{eqnarray}

\noindent where the two jackknife samples (for both $a$ and $b$) are:

@@

While the null test is useful for testing noise properties and the uniformity of a dataset, jackknives are useful in pinpointing which data subsets are contaminated by biases and which are not; in our toy model we see that the bias exists only in $\textbf{x}_{1}$ (Figure \ref{fig:toy_bias2}). If foreground or noise biases exist in a dataset, jackknives can tease them out and provide insight into possible sources. For example, if jackknives along the time-axis reveal a bias present at a certain LST, a likely explanation would be excess foreground emission from a radio source in the sky at that time. A jackknife test involving data before and after the application of a fringe-rate filter can reveal whether cross-talk noise bias is successfully suppressed with the filter, or if similar-shaped detections in both power spectras suggest otherwise. There are many other jackknife axes of which we will not go into detail here, including baseline, frequency, and polarization. Ultimately, an EoR detection should persist through them all and a clean measurement should exhibit noise-like null spectra.

>>>>

In the next three sections, we investigate the effects of weighting matrices on signal loss by experimenting with different matrices $\textbf{R}$ and examine their impact on the resulting power spectra estimate $\hat{\textbf{p}}$. Our ultimate goal in experimenting with weighting is to suppress foregrounds and investigate EoR losses associated with it. We note that we purposely take a thorough and pedagogical approach to describing the toy model examples given in the next few sections, for the sake of completeness. The specifics of how signal loss crops up in PAPER's analysis is later described in Section \ref{sec:Sigloss}.

@@

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.2cm 0.3cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss2.png}
	\caption{Eigenspectrum of $\hat{\textbf{C}}_{FG}$ (blue), $\hat{\textbf{C}}_{EoR}$ (red), and $\hat{\textbf{C}}_{FG+EoR}$ (green). The eigenspectrum of $\hat{\textbf{C}}_{FG}$ peaks at the zeroth eigenmode, due to the presence of only one sinusoid. These empirically estimated covariance matrices have eigenspectra that are different from that of a true $\textbf{C}$, meaning that some eigenmodes (shapes in the data) may be down-weighted more significantly than they should be, producing signal loss. Specifically, down-weighting the weak eigenmodes, where the EoR signal is greater than the FG signal, leads to loss.}
	\label{fig:toy_sigloss2}
\end{figure}

If the true covariance matrix $\textbf{C}$ of our data was known, then every single eigenvalue of $\textbf{C}$ would be representative of real fluctuations in the data. However, when using an estimated $\hat{\textbf{C}}$ that is derived from one particular data realization, its eigenvalues may differ from the truth. Said differently, shapes that may not exist (or have a weaker existence) in a true covariance may appear stronger in the estimated covariance. Hence, they will be down-weighted more than they should be.

In general, the strongest modes of $\hat{\textbf{C}}$ (highest eigenvalues) are expected to be dominated by bright foregrounds --- the most prominent shapes in the data. For these `strong' modes, where foregrounds outshine the EoR signal, $\hat{\textbf{C}} \sim \textbf{C}$ and down-weighting these modes is beneficial. This is demonstrated in the toy model by the successful suppression of the foreground mode, where our estimated covariance matrix identifies the sinusoid shape and assigns it the highest eigenvalue (the peak in Figure \ref{fig:toy_sigloss2}). In Figure \ref{fig:toy_sigloss12} we show the covariances of our toy model datasets along with inverse covariance weighted data. The foreground sinusoid is clearly visible in $\hat{\textbf{C}}_{FG}$ but effectively absent from $\hat{\textbf{C}}^{-1}_{FG}\textbf{x}_{FG}$.

@@

The danger of an empirically estimated covariance matrix comes from not being able to describe the `weak' eigenmodes of $\textbf{C}$ accurately, for which the EoR signal is brighter than foregrounds. If these weak eigenmodes of $\hat{\textbf{C}}$ characterize EoR modes, the EoR signal will be down-weighted (not as much as the strong foreground modes, but still down-weighted). This leads to the `overfitting' of noise (or `overfitting' of signal), which is signal loss.

Using what we've learned about the eigenspectrum, we can tweak it in a simple way to suppress foregrounds and yield zero signal loss. Recall that our toy model foreground is a sinusoid that can be perfectly described by a single eigenmode. We can project out the zeroth eigenmode (i.e. zero out all eigenmodes except the first one), thereby down-weighting the foregrounds perfectly and nothing else. Hence, we are not using information from all the weaker eigenmodes which carry the risk of describing EoR fluctuations. Altering $\hat{\textbf{C}}$ as such is one example of a regularization method, in which we are changing $\hat{\textbf{C}}$ in a way that flattens its eigenspectrum, making it more similar to that of $\textbf{I}$. The resulting power spectrum estimate for this case is shown in the right plot of Figure \ref{fig:toy_sigloss3}. In this case we recover EoR, demonstrating that if we can model our foregrounds perfectly (and disentangle them from other modes), we can down-weight them without signal loss. There are several other ways to regularize $\hat{\textbf{C}}$, and we will discuss some in Section \ref{sec:otherweight}.

@@

We have shown how signal loss can arise due to inaccurately characterizing weak eigenmodes (EoR dominated modes) with a data-estimated covariance. We will next show how this effect is exaggerated by reducing the total number of independent samples in a dataset.

@@

\begin{figure*}
	\centering
	\includegraphics[trim={1.5cm 0.2cm 1.5cm 0.3cm},clip,height=0.29\textwidth]{plots/toy_sigloss5.png}
	\includegraphics[trim={0.3cm 0.2cm 0.3cm 0.3cm},clip,height=0.29\textwidth]{plots/toy_sigloss15.png}
	\caption{Left: Our `fringe-rate filtered' (time-averaged) toy model dataset. We average every four samples together, yielding $25$ independent samples in time. Real parts are shown here. Right: Eigenspectrum of $\hat{\textbf{C}}_{FG+EoR}$, in the case of no fringe-rate filtering (solid green) and with fringe-rate filtering (dashed green). The dashed spectrum is steeper, resulting in more dramatic differences in weighting between eigenmodes. Consequently, if all eigenmodes are used in weighting it is possible to down-weight EoR modes more severely than for the non-averaged case.}
	\label{fig:toy_sigloss5}
\end{figure*}

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.2cm 0.3cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss7.png}
	\caption{Resulting power spectrum estimate for the fringe-rate filtered toy model simulation --- foregrounds only (blue), EoR only (red), and the weighted FG + EoR dataset (green). We use inverse covariance weighting where $\textbf{C}$ is computed from the data. There is a larger amount of signal loss than for the non-averaged data, a consequence of a steepened, weakly characterized eigenspectrum from using fewer independent modes in the data. The power spectrum estimate is also noisier, since most of the information used in constructing the estimate came from only a few eigenmodes (the weakest, EoR-dominated modes, which are down-weighted the least).}
	\label{fig:toy_sigloss7}
\end{figure}

@@

Using our toy model, we have seen that a sensitivity-driven analysis technique like fringe-rate filtering has trade-offs of signal loss and noisier estimates when using data-estimated covariance matrices. Longer integrations increase sensitivity but reduce the number of independent samples, resulting in poorly characterized, steep eigenspectra that can overfit signal greatly. We note that a fringe-rate filter does have a range of benefits, many described in \citet{parsons_et_al2016}, so it can still be advantageous to use one despite the trade-offs.

@@

In Section \ref{sec:toymodel} we showed one example (projecting the zeroth eigenmode) of how altering $\hat{\textbf{C}}$ can make the difference between zero and some signal loss, if we can distinguish between real eigenmodes in a true covariance matrix from those in an estimated one. We will now use our toy model to describe several other ways to tailor $\hat{\textbf{C}}$ in order to minimize signal loss. We choose four independent regularization methods to highlight in this section, which have been chosen due to their simplicity in implementation and straightforward interpretations. We illustrate the resulting power spectra and eigenspectra for the different cases in Figures \ref{fig:toy_sigloss8} and \ref{fig:toy_sigloss14}.

As a first test, we model the covariance matrix of EoR as a proof of concept that if perfect models are known, signal loss can be avoided. We know that our simulated EoR signal should have a covariance matrix that mimics the identity matrix, with its variance encoded along the diagonal. We model $\textbf{C}_{EoR}$ as such, instead of computing it based on $\textbf{x}_{EoR}$ itself. Next, we add $\textbf{C}_{EoR} + \hat{\textbf{C}}_{FG}$ (where $\hat{\textbf{C}}_{FG} = \langle\textbf{x}_{FG}\textbf{x}_{FG}^{\dagger}\rangle$) to obtain a final $\hat{\textbf{C}}$ to use in weighting. In Figure \ref{fig:toy_sigloss8} (upper left), we see that there is negligible signal loss. This is because by modeling $\textbf{C}_{EoR}$, we avoid over-fitting fluctuations in the data that our model doesn't know about (but an empirically derived $\hat{\textbf{C}}$ would). This is evident when comparing the green and red curves in Figure \ref{fig:toy_sigloss14}. We don't see perfect EoR recovery though, as the green and red curves vary slightly in Figure \ref{fig:toy_sigloss8}. This deviation is a consequence of the difference between the true and modeled covariance matrices. In practice such a weighting option is not feasible, as we do not know $\textbf{C}_{EoR}$.

@@

The second panel (top right) in Figure \ref{fig:toy_sigloss8} uses a regularization method of setting $\hat{\textbf{C}} = \hat{\textbf{C}} + \gamma\textbf{I}$, where $i$$=$$j$$=$$20$ (number of frequencies) and $\gamma = 5$ (an arbitrary strength of $\textbf{I}$ for the purpose of this toy model). By adding the identity matrix, element-wise, we are weighting the diagonal elements of the matrix more heavily than those off-diagonal, thereby flattening out its eigenspectrum. If all modes are given similar weights, we avoid down-weighting EoR modes more than others.

@@

Our second major $21$ cm power spectrum theme is error estimation, as we desire robust methods for determining accurate confidence intervals for our measurements. Two popular ways of empirically calculating errors on a power spectrum measurement are calculating the variance of a dataset, and computing a theoretical error estimate based on an instrument's system temperature and observational parameters. In a perfect world, both methods would match up. However, in practice the two don't always agree due to a number of factors, including time, frequency, and antenna-dependent noise and non-uniform weightings.

For PAPER's analysis, we choose to compute error bars on our measurements that have been derived from its inherent variance. A common technique used to do this is bootstrapping. For pedagogical purposes, we first define the technique of bootstrapping and then illustrate some of its pitfalls through toy models.

Bootstrapping uses sampling with replacement to estimate a posterior distribution. For example, measurements (like power spectra) can be made from different samples of data. Averaging each sample set gives us a realization of the (power spectrum) answer. Realizations are correlated with each other to a degree set by the fraction of sampled points that are held in common between them. Hence, through the process of resampling and averaging along different axes, we can estimate error bars for our results which represent the underlying distribution of values that are allowed by our measurements.

Suppose we have $N$ different measurements targeting the same quantity ($N$ power spectrum measurements). Bootstrapping means that we form $N_{boot}$ (often a large number) bootstraps, where each bootstrap is a random selection of the $N$ measurements. Bootstraps each have dimensions of $N$, and the values populated into each bootstrap are drawn from the original set of measurements with replacement (every $N$th slot is filled randomly for each bootstrap). Next we take the mean of each bootstrap to collapse it from an array of length $N$ to a single number (we are interested in the mean statistic here, but any function of interest can be applied to each bootstrap as long as it's the same function for each one). The error (on the mean) is then computed as the standard deviation across all bootstraps.

We must be careful in distinguishing $N_{boot}$, the number of bootstraps, from $N$, the number of samples, or elements, or values, that comprise a bootstrap. In the toy models presented in this section, $N_{boot}$ is typically large, and the standard deviation across bootstraps (the error we are computing) converges for large $N_{boot}$. It is the value of $N$ that we must be careful in setting, as described in the rest of this section.

For our toy model, suppose we have a Gaussian random signal dataset of length $N=1000$ and unity variance (zero mean). This could represent $1000$ power spectrum measurements, for which we are interested in its error. We predict that the error on the mean should obey $1/\sqrt{N}$, where $N$ is the number of samples.

We next form $500$ bootstraps ($N_{boot} = 500$). To create each bootstrap, we draw $N$ samples, with replacement, of the original data, and take the mean over the $N$ samples. The standard deviation over the $500$ bootstraps gives an error estimate for our dataset. This error is indicated by the grey star in Figure \ref{fig:toy_error1} and matches our theoretical prediction (green).

@@

This example highlights the importance of understanding how analysis techniques (e.g. fringe-rate filtering) can affect a common statistical procedure like bootstrapping. Bootstrapping as a means of estimating power spectrum errors from real fringe-rate filtered data requires knowledge of the number of independent samples, which is not always a trivial task. For example, computing the effective number of independent samples of fringe-rate filtered data is not as simple as counting the number of averages performed. Down-sampling a time-averaged signal is straightforward using a boxcar average, but non-trivial with a more complicated convolution function that has long tails. Hence, we do not recommend bootstrapping unless the number of independent samples along the axis that is being re-sampled is well-determined.

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.3cm 0.3cm 0.3cm},width=\columnwidth]{plots/toy_error1.png}
	\caption{Error estimation produced by bootstrapping as a function of the number of elements drawn per bootstrap when sampling with replacement. The star represents the standard deviation of $500$ bootstraps, each created by drawing $1000$ elements (with replacement) from a length $1000$ array of a Gaussian random signal. The black points correspond to time-averaged data (correlated data) which has $100$ independent samples. They illustrate how errors can be underestimated if drawing more elements than there are independent samples in the data. The estimated errors match up with the theoretical prediction only at $N=100$.}
	\label{fig:toy_error1}
\end{figure}

We will now discuss a second subtle feature of bootstrapping that fails to maximize the sensitivity of a dataset. Suppose we have $5$ independent measurements of the sky (from $5$ different baselines, for example). A bootstrap then consists of $5$ measurements that are drawn randomly with replacement from the original set. If all $5$ spaces are filled randomly, there is a high probability that some measurements will be repeated in the bootstrap because they are drawn more than once. The bootstrap may therefore consist of only $3$ or $4$ independent measurements --- a number smaller than the total number of samples. In fact, the probability of drawing $5$ completely independent measurements is less than $4\%$.

@@

Therefore, we use a slightly modified bootstrapping method. For each bootstrap, we draw $N-1=4$ samples without replacement and draw the last slot randomly with replacement. In doing so there is a small chance a bootstrap will consist of $5$ independent measurements, but even with $4$ our sensitivity is nearly maximized. %One may wonder whether this change is still a legitimate way of error estimating since random sampling only occurs for one value in a bootstrap. However, as long as the number of possible variations is greater than the number of bootstraps we perform, it is a valid way to uncover the inherent variability in a dataset.

@@

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.3cm 0.3cm 0.3cm},width=\columnwidth]{plots/toy_error2.png}
	\caption{The fraction of independent elements in one bootstrap sample as a function of the total number of elements in a dataset (i.e. if $4$ out of $5$ elements are independent, this would yield a fraction of $0.8$). A fraction of $1.0$ means that all elements in the bootstrap sample are independent, and therefore sensitivity is maximized (alternately, this axis can be thought of as power spectrum sensitivity, where sensitivity increases moving upwards along the y-axis). Two bootstrapping methods are shown here --- sampling all elements with replacement (grey) and sampling only the last element with replacement (black). The former fails to maximize sensitivity since some elements end up being repeated due to random sampling.} %The `outlier' points lying at a fraction of $1.0$ at the far left of the plot occur in the off-chance that the last spot in the sample happens to be the one missing element instead of a repeated element --- this is more likely to happen with a small total number of elements but also results in a more dramatically different fraction as compared to the case of a repeated element in the last spot, hence why these points appear like outliers.}
	\label{fig:toy_error2}
\end{figure}

@@

However, this isolation is not perfect.  In deep measurements, power spectrum measurements at $k_{\parallel}$ values beyond the length of a baseline are often still contaminated at a low level. This leakage, especially at low $k$'s, can be attributed to convolution kernels associated with Fourier-transforming visibilities into delay-space. In other words, smooth-spectrum foregrounds appear as $\delta$-functions in delay-space, convolved by the Fourier transform of the source spectrum and the antenna response, both of which could smear out the foregrounds and cause leakage outside the wedge.

@@

Confirming foreground detections at higher $k$'s is more difficult. In the next section (Section \ref{sec:JackknifeOverview}), we will present some tests that can help distinguish these excesses from that of EoR.

@@

@@

We now approach the difficult task of tracing excesses to foreground, noise, and EoR biases through a discussion of useful jackknife tests. The jackknife is a resampling technique in which a statistic (i.e. power spectrum) is computed in subsets of the data. These subsets are then compared to reveal systematics. In this section we define two main tests --- the null test and the traditional jackknife --- and explain how a power spectrum detection must pass each. We then highlight how these tests can be used to help distinguish between different sources of bias.

\begin{itemize}
\item \textbf{Null Test}: A null test is a type of jackknife test that removes the astronomical signal from data in order to investigate underlying systematics. Several null tests on data are described in \citet{keating_et_al2016}. For example, one can divide data into two subsets by separating odd and even Julian dates, or the first half of the observing season from the second. Subtracting the two removes signal that is common to both subsets, including foregrounds and EoR. The resulting power spectrum should be consistent with thermal noise estimates; if it is not, it suggests the presence of a systematic that differs from one of the data subsets to the other (i.e. doesn't get subtracted perfectly). 
\item \textbf{Traditional Jackknife}: In a more broad sense, it is important to perform many jackknife tests in order to instill confidence in a final result. A stable result must be steadfast throughout all jackknives no matter how the data is sliced. Jackknives can be taken along several different axes --- for example, one could start with a full dataset, and compute a new power spectrum every time as a day of data is removed, or a baseline is removed. This type of jackknife would reveal bias present only at certain LSTs (such as a foreground source), for example, or misbehaving baselines.
\end{itemize}

@@

\begin{equation}
\label{eq:bias1}
\textbf{x}_{a} = \textbf{n}_{a} + \textbf{e} + \textbf{fg}
\end{equation}

\begin{equation}
\label{eq:bias2}
\textbf{x}_{b} = \textbf{n}_{b} + \textbf{e} + \textbf{fg}.
\end{equation}

\noindent The two jackknife samples are $\textbf{x}_{1}$ and $\textbf{x}_{2}$, representing jackknives in time. These can be written (for both measurements $a$ and $b$) as:

@@

While the null test is useful for testing noise properties and the uniformity of a dataset, jackknives are useful in pinpointing which data subsets are contaminated by biases and which are not; in our toy model we see that the bias exists only in $\textbf{x}_{1}$ (Figure \ref{fig:toy_bias2}). If foreground or noise biases exist in a dataset, jackknives can tease them out and provide insight into possible sources. For example, if jackknives along the time-axis reveal a bias present at a certain LST, a likely explanation would be excess foreground emission from a radio source in the sky at that time. A jackknife test involving data before and after the application of a fringe-rate filter can reveal whether cross-talk noise bias is successfully suppressed with the filter, or if similar-shaped detections in both power spectra suggest otherwise. There are many other jackknife axes of which we will not go into detail here, including baseline, frequency, and polarization. Ultimately, an EoR detection should persist through them all and a clean measurement should exhibit noise-like null spectra.
